---
title: 'Week 9: Lab - Using aRules on the Titanic dataset '
author: 'Tristan Wijnmaalen'
date: '03/12/23'
output: html_document
---

---

# Instructions
Use the Titanic dataset to explore descriptive statistics, functions, and
association rules. Download the titanic dataset `titanic.raw.rdata` from one
of two locations:

- https://github.com/ethen8181/machine-learning/tree/master/association_rule/R
- https://sites.google.com/a/rdatamining.com/www/data/titanic.raw.rdata?attredirects=1
 
Note that it is not a cvs file, but rather an RData workspace. So, to load the 
data (assuming you saved it to the project's data folder), one would do:

``load("data/titanic.raw.rdata")``

You need to look at titanic.raw (the name of the R dataset)

``t <- titanic.raw``

Now that you have the datafile, do some descriptive statistics, getting some 
extra practice using R. Take the Quiz for Step 1 and 2. However your submission 
**MUST** include the process of you calculating the following values in Step 1 and 2.

---

```{r setup, message = FALSE}
# Add your library below.
library(arules)


```

# Step 0 - Load the data
Using the instructions above, load the dataset and save it as `t`.

```{r, "Step 0"}
# Write your code below.
load("data/titanic.raw.rdata")
t <- titanic.raw


```
---


# Step 1 - Descriptive stats (0.5 point for each answer)

1. Compute the percentage of people that survived.
2. Compute the percentage of people that were children.
3. Compute the percentage of people that were female.
4. Finally, compute the percentage of people that were in first class.

```{r, "Step 1"}
# Write your code below.
survived <- as.character(t$Survived)
age <- as.character(t$Age)
class <- as.character(t$Class)
sex <- as.character(t$Sex)

transactions <- read.transactions(textConnection(survived))
itemFrequency(transactions)["Yes"]

transactions <- read.transactions(textConnection(age))
itemFrequency(transactions)["Child"]

transactions <- read.transactions(textConnection(sex))
itemFrequency(transactions)["Female"]

transactions <- read.transactions(textConnection(class))
itemFrequency(transactions)["1st"]
```

---

# Step 2 - More descriptive stats (0.5 point for each answer)


1. What percentage of children survived? Your answer should be written such as # 13.75% of children survived
2. What percentage of female survived?
3. What percentage of first-class people survived?
4. What percentage of third-class people survived?

```{r, "Step 2"}
# Write your code below.
n.child.survived <- nrow(subset(t, Age == "Child" & Survived == "Yes"))
cat(round(n.child.survived/sum(t$Age == "Child")*100,2), "% of children survived\n", sep = "")

n.female.survived <- nrow(subset(t, Sex == "Female" & Survived == "Yes"))
cat(round(n.female.survived/sum(t$Sex == "Female")*100,2), "% of female survived\n", sep = "")

n.class1.survived <- nrow(subset(t, Class == "1st" & Survived == "Yes"))
cat(round(n.class1.survived/sum(t$Class == "1st")*100,2), "% of first-class survived\n", sep = "")

n.class3.survived <- nrow(subset(t, Class == "3rd" & Survived == "Yes"))
cat(round(n.class3.survived/sum(t$Class == "3rd")*100,2), "% of third-class survived\n", sep = "")
```

---

# Step 3 - Writing a function (0.5 point for each answer)

## Step 3.1 - Function 1
Write a function that returns a new dataframe of people that satisfies the 
specified criteria of sex, age, class and survived as parameters. I'm giving you 
the answer for this question:

```
myfunction1 <- function(a,b,c,d){
df1 <- t[t$Class == a,] # filter the data that satisfied the criteria that "Class" = a
df2 <- df1[df1$Sex == b,] # filter the data that satisfied the criteria that "Sex" = b
df3 <- df2[df2$Age == c,] # filter the data that satisfied the criteria that "Age" = c
df4 <- df3[df3$Survived == d,] # filter the data that satisfied the criteria that "Survived" = d
return(df4)}
# test the function with a sample data
myfunction1("1st","Female","Adult","No")
```

```{r, "Step 3.1"}
# Write your code below.
myfunction1 <- function(a,b,c,d){
df1 <- t[t$Class == a,] # filter the data that satisfied the criteria that "Class" = a
df2 <- df1[df1$Sex == b,] # filter the data that satisfied the criteria that "Sex" = b
df3 <- df2[df2$Age == c,] # filter the data that satisfied the criteria that "Age" = c
df4 <- df3[df3$Survived == d,] # filter the data that satisfied the criteria that "Survived" = d
return(df4)}
# test the function with a sample data
myfunction1("1st","Female","Adult","No")


```

## Step 3.2 - Function 2
Write a function, using the previous function, that calculates the percentage 
(who lives, who dies) for a specified (parameters) of class, gender and age 
considering the entire number of data. The function passes four arguments. 
Include the following code properly in your function by improvising names of objects.

```
p <- nrow(df)/nrow(t) # calculate the percentage
```

```{r, "Step 3.2"}
# Write your code below.
myfunction2 <- function(a,b,c,d){
  df1 <- t[t$Class == a,]
  df2 <- df1[df1$Sex == b,]
  df3 <- df2[df2$Age == c,]
  df4 <- df3[df3$Survived == d,]
  cat(round(nrow(df4)/nrow(df3)*100,2), "% of people with these parameters survived\n", sep = "")
  dfsurvivors <- t[t$Survived == "Yes",]
  cat(round(nrow(df4)/nrow(dfsurvivors)*100,2), "% of survivors have these parameters\n", sep = "")
  return(cat(round(nrow(df4)/nrow(t)*100,2), "% of people who boarded have these parameters\n", sep = ""))
}
myfunction2("1st","Female","Child","Yes")

```

## Step 3.3 - Use the function (male)
Use the function to compare age and third-class male survival rates. 

```{r, "Step 3.3"}
# Write your code below.
myfunction2("3rd","Male","Adult","Yes")
myfunction2("3rd","Male","Child","Yes")

```
People in which category are more likely to survive? 

> The second category, children.


## Step 3.4 - Use the function (female)
Use the function to compare age and first-class female survival rates.

```{r, "Step 3.4"}
# Write your code below.
myfunction2("1st","Female","Adult","Yes")
myfunction2("1st","Female","Child","Yes")

```
People in which category are more likely to survive? 

> Children, but there was only one first-class female child onboard. 

---

# Step 4 - Use aRules (0.5 point for each answer)

1. Use aRules to calculate some rules (clusters) for the titanic dataset.
2. Visualize the results.
3. Pick the three most interesting and useful rules. Explain these rules using natural language. **Answer this in the space provided below.**
4. How does this compare to the descriptive analysis we did on the same data set? 
Think critically. What was possible using one method that was not possible using 
the other method? **Answer this in the space provided below.**

```{r, "Step 4.1"}
# Write your code below.
transactions <- as(t, "transactions")
rules <- apriori(transactions, parameter = list(support = 0.005, confidence = 0.5),
                 appearance = list(default="lhs", rhs=("Survived=Yes")))
summary(rules)
inspect(rules)
rules.sortedlift <- sort(subset(rules, lift > 2), by = "lift", decreasing = TRUE)
inspect(rules.sortedlift[1:10])

```

**Answer part 3 and 4 below.**

> The highest lift rule indicating that 2nd class children survived the most is interesting, and I'm wondering if it's like this because there weren't many children in 1st class. The 2nd highest lift rule which is fully tied with the first specifies female children having the highest survival rate. Finally, the 3rd rule puts 1st class women of any age at third most likely to survive. 
This second method seems to more fairly compare groups of different sizes, specifically the singular first class female child is not at the top of the list even though they resulted in that category having a 100% survival rate.


